---
title: "Identifying Chinese Children with Dyslexia via Machine Learning"

---
## Setup MLP Python Environment & Error Prevention
```{r}
# setup-python-env
library(reticulate)

# Define the name of the virtual environment for the MLP model
env_name <- "r-mlp"

# 1) Suggest users to set their own Python path if necessary 
# (Uncomment and modify the path below if R cannot find Python automatically)
# use_python("C:/path/to/your/python.exe", required = TRUE)

# 2) Create the virtual environment if it does not exist
if (!virtualenv_exists(env_name)) {
  virtualenv_create(env_name)
}

# 3) Activate the virtual environment
use_virtualenv(env_name, required = TRUE)

# 4) Install TensorFlow and Keras into the designated environment
if (!reticulate::py_module_available("tensorflow")) {
  cat("TensorFlow not found. Installing now...\n")
  tensorflow::install_tensorflow(method = "virtualenv",
                                 envname = env_name,
                                 version = "2.15.0")
}

# 5) Verify the current Python configuration
reticulate::py_config()
```



## Load Packages, Paths, and Seeds
```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(purrr)
  library(stringr)
  library(tibble)
  library(ggplot2)
  library(haven)
  library(caret)
  library(pROC)
  library(recipes)
  library(themis)
  library(glmnet)
  library(vip)
  library(yardstick)
  library(rsample)
  library(xgboost)

  library(fastshap)
  library(shapviz)
  library(keras)
  library(iml)
})

options(dplyr.summarise.inform = FALSE)  

set.seed(111) 

# =====================================================================
# Environment Setup & Working Directory
# =====================================================================
# Note for reproducibility: 
# Please ensure that this .Rmd script and all associated .rds data files 
# are stored in the same folder on your local machine. 
# 
# Before running the code, set your R working directory to that folder.
# (In RStudio, you can do this via: Session -> Set Working Directory -> To Source File Location)

```


# load database
```{r}
RD_4ML <- readRDS("dataset_anonymized.rds")
```


# set model var
```{r}
predictor_vars_1 <- c( # Include all variables in the analysis
  "gender", "Grade",
  "LxclScor", "FlldScor",
  "cnfgrtnLR", "cnfgrtnUD", 
  "RdcNoMuti","RdcNodub", 
  "trspT","trspV", 
  "SRngA","SRngH", 
  "PType_Onset","PType_Rime","PType_Tone","PType_None","PType_Mark",
  "PType_1of3","PType_2of3","PType_3of3",
  "LxclScor", "FlldScor",  
  "MMCScor","MMPScor","MMFScor", 
  "COnstScor","CRimScor","CTnScor", 
  "POnstScor","PRimScor","PTnScor" 
)

predictor_vars_2 <- c( # Whole-character level - Phonology
  "LxclScor",
  "MMCScor", 
  "COnstScor","CRimScor","CTnScor",
  "PType_Onset","PType_Rime","PType_Tone"
)


predictor_vars_3 <- c( # Whole-character level - Orthography
  "cnfgrtnLR", "cnfgrtnUD", 
  "RdcNoMuti","RdcNodub"
)

predictor_vars_4 <- c( # Whole-character level - Semantics
  "SRngA","SRngH"
)

predictor_vars_5 <- c( # Sub-lexical level - Phonology
  "MMPScor","POnstScor","PRimScor","PTnScor", 
   "PType_1of3","PType_2of3","PType_3of3","PType_None"
)

predictor_vars_6 <- c( # Sub-lexical level - Orthography
  "PType_Mark","MMFScor"
)

predictor_vars_7 <- c( # Sub-lexical level - Semantics
  "trspT","trspV"
)

```



# t test
```{r}
vars_2_4T <- setdiff(predictor_vars_1, c("gender", "Grade", "DD"))
RD_4ML_T <- RD_4ML %>%
  dplyr::select(all_of(vars_2_4T), DD)

num_vars <- names(dplyr::select(RD_4ML_T, where(is.numeric), -DD))

RD_4ML_T <- RD_4ML_T %>%
  dplyr::mutate(across(all_of(num_vars), ~ as.numeric(scale(.))))

T_results <- list()
for (var in vars_2_4T) {
  if (!var %in% names(RD_4ML_T)) next
  if (!is.numeric(RD_4ML_T[[var]])) next
  formula <- as.formula(paste(var, "~ DD"))
  t_test_raw <- RD_4ML_T %>%
    rstatix::t_test(formula, var.equal = TRUE, detailed = TRUE)
  
  t_test_result <- t_test_raw %>%
    dplyr::transmute(
      t  = statistic,
      df = df,
      p_num = as.numeric(p),
      p_sci = format(p, scientific = TRUE, digits = 3),
      p_value = dplyr::if_else(p < 1e-4, "<0.0001", sprintf("%.4f", p))
    )

  descriptives <- RD_4ML_T %>%
    dplyr::group_by(DD) %>%
    dplyr::summarise(
      n = dplyr::n(),
      mean = mean(.data[[var]], na.rm = TRUE),
      sd = sd(.data[[var]], na.rm = TRUE),
      .groups = "drop"
    ) %>%
    tidyr::pivot_wider(
      names_from  = DD,
      values_from = c(n, mean, sd),
      names_sep   = "_"
    )

  # Cohen's d
  cohen_d_result <- RD_4ML_T %>%
    rstatix::cohens_d(formula, ci = TRUE) %>%
    dplyr::select(effsize, conf.low, conf.high) %>%
    dplyr::rename(cohen_d = effsize, ci_low = conf.low, ci_high = conf.high)

  combined_result <- dplyr::bind_cols(descriptives, t_test_result, cohen_d_result)

  T_results[[var]] <- combined_result
}

final_results_table <- dplyr::bind_rows(T_results, .id = "Variable")
print(final_results_table)


final_table_formatted <- final_results_table %>%
  dplyr::mutate(
    `DD (n=496) M (SD)` = sprintf("%.2f (%.2f)", mean_Dyslexia, sd_Dyslexia),
    `TD (n=6770) M (SD)` = sprintf("%.2f (%.2f)", mean_Normal, sd_Normal),

    t = sprintf("%.2f", t),
    
    # "d [ci_low, ci_high]"
    `Cohen's d [95% CI]` = sprintf("%.2f [%.2f, %.2f]", cohen_d, ci_low, ci_high)
  ) %>%
  dplyr::select(
    Variable,
    `DD (n=496) M (SD)`,
    `TD (n=6770) M (SD)`,
    t,
    p_value,
    `Cohen's d [95% CI]`
  )
final_table_formatted
```


# data seperating 
```{r}
set.seed(111)
train_idx <- createDataPartition(RD_4ML$DD, p = 0.8, list = FALSE)
train_full <- RD_4ML[train_idx, ]
test_full <- RD_4ML[-train_idx, ]
```



## create sub-dataset
```{r}
# database for predictor_vars_1
RD_4ML_1_train <- train_full %>%
  select(all_of(c("DD", predictor_vars_1)))
RD_4ML_1_test <- test_full %>%
  select(all_of(c("DD", predictor_vars_1)))

# database for predictor_vars_2
RD_4ML_2_train <- train_full %>% 
  select(all_of(c("DD", predictor_vars_2)))
RD_4ML_2_test <- test_full %>% 
  select(all_of(c("DD", predictor_vars_2)))

# database for predictor_vars_3
RD_4ML_3_train <- train_full %>% 
  select(all_of(c("DD", predictor_vars_3)))
RD_4ML_3_test <- test_full %>% 
  select(all_of(c("DD", predictor_vars_3)))

# database for predictor_vars_4
RD_4ML_4_train <- train_full %>% 
  select(all_of(c("DD", predictor_vars_4)))
RD_4ML_4_test <- test_full %>% 
  select(all_of(c("DD", predictor_vars_4)))

# database for predictor_vars_5
RD_4ML_5_train <- train_full %>% 
  select(all_of(c("DD", predictor_vars_5)))
RD_4ML_5_test <- test_full %>% 
  select(all_of(c("DD", predictor_vars_5)))

# database for predictor_vars_6
RD_4ML_6_train <- train_full %>% 
  select(all_of(c("DD", predictor_vars_6)))
RD_4ML_6_test <- test_full %>% 
  select(all_of(c("DD", predictor_vars_6)))

# 数据库6：predictor_vars_7
RD_4ML_7_train <- train_full %>% 
  select(all_of(c("DD", predictor_vars_7)))
RD_4ML_7_test <- test_full %>% 
  select(all_of(c("DD", predictor_vars_7)))

```

# tools
```{r}

# Threshold-based binary classification counts and derived metrics (avoids yardstick zero-event warnings)
.bin_counts <- function(truth, prob, thr){
  pred <- factor(ifelse(prob >= thr, "Dyslexia", "Normal"),
                 levels = c("Dyslexia","Normal"))
  tp <- sum(truth=="Dyslexia" & pred=="Dyslexia", na.rm=TRUE)
  fp <- sum(truth=="Normal"   & pred=="Dyslexia", na.rm=TRUE)
  tn <- sum(truth=="Normal"   & pred=="Normal",   na.rm=TRUE)
  fn <- sum(truth=="Dyslexia" & pred=="Normal",   na.rm=TRUE)
  list(tp=tp, fp=fp, tn=tn, fn=fn,
       pred=pred)
}
.metrics_from_counts <- function(cnt){
  n  <- cnt$tp + cnt$fp + cnt$tn + cnt$fn
  se <- ifelse(cnt$tp+cnt$fn>0, cnt$tp/(cnt$tp+cnt$fn), NA_real_)   # Sensitivity
  sp <- ifelse(cnt$tn+cnt$fp>0, cnt$tn/(cnt$tn+cnt$fp), NA_real_)   # Specificity
  pr <- ifelse(cnt$tp+cnt$fp>0, cnt$tp/(cnt$tp+cnt$fp), NA_real_)   # Precision
  acc <- (cnt$tp + cnt$tn)/n
  f1  <- ifelse(pr+se>0, 2*pr*se/(pr+se), NA_real_)
  f2  <- ifelse(4*pr+se>0, 5*pr*se/(4*pr+se), NA_real_)
  tibble::tibble(Accuracy=acc, Sensitivity=se, Specificity=sp, Precision=pr, F1=f1, F2=f2)
}

# Generate threshold table
.pr_table <- function(truth, prob, ths = seq(0.01,0.99,by=0.01)){
  m <- lapply(ths, function(t){
    cnt <- .bin_counts(truth, prob, t)
    c(threshold=t, .metrics_from_counts(cnt))
  })
  dplyr::bind_rows(lapply(m, as.data.frame))
}

# Priority: Maximize Sensitivity, then Specificity, then F1 score.
pick_threshold_sen_first <- function(truth, prob, ths = seq(0.01,0.99,by=0.01)){
  tab <- .pr_table(truth, prob, ths)
  mx_se <- max(tab$Sensitivity, na.rm = TRUE)
  cand  <- tab[which(tab$Sensitivity >= mx_se - 1e-12), , drop=FALSE]
  mx_sp <- max(cand$Specificity, na.rm = TRUE)
  cand2 <- cand[which(cand$Specificity >= mx_sp - 1e-12), , drop=FALSE]
  idx   <- which.max(cand2$F1)
  list(th = cand2$threshold[idx], table = tab)
}

# Calculate common metrics for a given threshold：
.metrics_at_threshold <- function(truth, prob, th){
  pred <- factor(ifelse(prob >= th, "Dyslexia","Normal"),
                 levels = c("Dyslexia","Normal"))
  tp <- sum(truth=="Dyslexia" & pred=="Dyslexia")
  fp <- sum(truth=="Normal"   & pred=="Dyslexia")
  tn <- sum(truth=="Normal"   & pred=="Normal")
  fn <- sum(truth=="Dyslexia" & pred=="Normal")
  prec <- ifelse(tp+fp>0, tp/(tp+fp), NA_real_)
  rec  <- ifelse(tp+fn>0, tp/(tp+fn), NA_real_)            #  Sensitivity
  spec <- ifelse(tn+fp>0, tn/(tn+fp), NA_real_)
  acc  <- (tp+tn)/(tp+fp+tn+fn)
  f1   <- ifelse(prec+rec>0, 2*prec*rec/(prec+rec), NA_real_)
  f2   <- ifelse(4*prec+rec>0, 5*prec*rec/(4*prec+rec), NA_real_)
  list(Accuracy=acc, Sensitivity=rec, Specificity=spec,
       Precision=prec, F1=f1, F2=f2)
}

threshold_select_composite <- function(truth, prob,
                                       method = c("youden","closest_topleft"),
                                       tie_breaker = c("F1","Precision")){
  method      <- match.arg(method)
  tie_breaker <- match.arg(tie_breaker)

  roc_obj <- pROC::roc(truth, prob, levels = c("Dyslexia","Normal"), direction = ">")
  tab <- as.data.frame(pROC::coords(
    roc_obj, x = "all",
    ret = c("threshold","sensitivity","specificity","accuracy","ppv"), 
    transpose = TRUE
  ))
  names(tab) <- c("thr","se","sp","acc","ppv")

  tab$se  <- as.numeric(tab$se)
  tab$sp  <- as.numeric(tab$sp)
  tab$acc <- as.numeric(tab$acc)
  tab$ppv <- as.numeric(tab$ppv)
  tab$J   <- tab$se + tab$sp - 1
  tab$dist <- sqrt((1 - tab$se)^2 + (1 - tab$sp)^2)
  tab$F1  <- ifelse(tab$ppv + tab$se > 0, 2*tab$ppv*tab$se/(tab$ppv + tab$se), NA_real_)

  if (method == "youden"){
    idx <- which(tab$J == max(tab$J, na.rm = TRUE))
    # # Break ties using F1 or Precision
    if (length(idx) > 1){
      pick <- tab[idx, ]
      ord  <- order(pick$F1, pick$ppv, decreasing = TRUE)
      idx  <- idx[ord][1]
    } else idx <- idx[1]
  } else { # closest_topleft
    idx <- which(tab$dist == min(tab$dist, na.rm = TRUE))
    if (length(idx) > 1){
      pick <- tab[idx, ]
      ord  <- if (tie_breaker == "F1") order(pick$F1, pick$ppv, decreasing = TRUE)
              else                     order(pick$ppv, pick$F1, decreasing = TRUE)
      idx  <- idx[ord][1]
    } else idx <- idx[1]
  }
  th <- tab$thr[idx]
  list(th = th, table = tab, pick = idx)
}

make_recipe <- function(df,
                        smote_ratio = 0.6,
                        id_cols     = c(".row_id")) {
  if (!is.factor(df$DD)) {
    df$DD <- factor(df$DD,
                    levels = c("Dyslexia", "Normal")) 
  } else {
    df$DD <- factor(df$DD, levels = c("Dyslexia", "Normal"))
  }

  rec <- recipes::recipe(DD ~ ., data = df)
  keep_ids <- intersect(id_cols, names(df))
  if (length(keep_ids) > 0) {
    rec <- recipes::update_role(rec, tidyselect::all_of(keep_ids), new_role = "id")
  }

  rec <- rec %>%
    recipes::step_string2factor(recipes::all_nominal_predictors()) %>%  # Convert strings to factors
    recipes::step_novel(recipes::all_nominal_predictors()) %>%          # Mark novel levels in test set as "new"
    recipes::step_dummy(recipes::all_nominal_predictors(), one_hot = TRUE) %>%  # One-hot encodin
    recipes::step_zv(recipes::all_predictors()) %>%                    # Remove zero-variance predictors
    recipes::step_normalize(recipes::all_numeric_predictors()) %>%     # Standardize (mean = 0, SD = 1)
    themis::step_smote(DD, over_ratio = smote_ratio, seed = 111, skip = TRUE)  # SMOTE
  rec
}

 # 
make_ConSens_summary <- function(spec_floor = 0.80,
                                 acc_floor  = 0.80,
                                 ths = seq(0.01, 0.99, by = 0.01)) {
  pr_table_local <- function(truth, prob, ths) {
    lev1 <- levels(truth)[1]; lev2 <- levels(truth)[2]
    out <- lapply(ths, function(t){
      pred <- factor(ifelse(prob >= t, lev1, lev2), levels = c(lev1, lev2))
      tp <- sum(truth==lev1 & pred==lev1); fp <- sum(truth==lev2 & pred==lev1)
      tn <- sum(truth==lev2 & pred==lev2); fn <- sum(truth==lev1 & pred==lev2)
      se  <- ifelse(tp+fn>0, tp/(tp+fn), 0)
      sp  <- ifelse(tn+fp>0, tn/(tn+fp), 0)
      acc <- ifelse(tp+fp+tn+fn>0, (tp+tn)/(tp+fp+tn+fn), 0)
      data.frame(threshold=t, Sensitivity=se, Specificity=sp, Accuracy=acc)
    })
    do.call(rbind, out)
  }
  function(data, lev = NULL, model = NULL){
    if (is.null(lev) || length(lev)!=2) return(c(ConSens=0))
    if (!lev[1] %in% names(data) || !"obs" %in% names(data)) return(c(ConSens=0))
    truth <- factor(data$obs, levels = lev)
    prob  <- suppressWarnings(as.numeric(data[[lev[1]]]))
    if (!any(is.finite(prob))) return(c(ConSens=0))
    tab <- pr_table_local(truth, prob, ths)
    ok  <- with(tab, Specificity>=spec_floor & Accuracy>=acc_floor &
                      is.finite(Sensitivity) & is.finite(Specificity) & is.finite(Accuracy))
    if (!any(ok)) return(c(ConSens=0))
    c(ConSens = max(tab$Sensitivity[ok], na.rm = TRUE))
  }
}

pick_threshold_constrained <- function(truth, prob,
                                       ths, target_sens_val, min_spec_val, min_acc_val){
  tab  <- .pr_table(truth, prob, ths)
  feas <- subset(tab, Specificity >= min_spec_val & Accuracy >= min_acc_val)
  feas90 <- subset(feas, Sensitivity >= target_sens_val)
  if (nrow(feas90)){
    J <- with(feas90, Sensitivity + Specificity - 1)
    best <- feas90[order(-J, -feas90$F1), ][1,,drop=FALSE]
    return(list(th=best$threshold, table=tab, best_row=best, met_all=TRUE))
  }
  if (nrow(feas)){
    J <- with(feas, Sensitivity + Specificity - 1)
    best <- feas[order(-feas$Sensitivity, -J, -feas$F1), ][1,,drop=FALSE]
    return(list(th=best$threshold, table=tab, best_row=best, met_all=FALSE))
  }
  J <- with(tab, Sensitivity + Specificity - 1)
  best <- tab[order(-J, -tab$F1), ][1,,drop=FALSE]
  list(th=best$threshold, table=tab, best_row=best, met_all=FALSE)
}

get_result_obj <- function(prefix, i, env = parent.frame()){
  get0(sprintf("%s_15sd_results_%d", prefix, i), envir = env, ifnotfound = NULL)
}

fetch_probs_safe <- function(obj){
  if (is.null(obj)) return(NULL)
  if (!is.null(obj$probs_test)) return(as.numeric(obj$probs_test))
  NULL
}

.gen_weight_grid <- function(M, step = 0.10, tol = 1e-9){
  vals <- seq(0, 1, by = step)
  rec <- function(k, rest){
    if (k == M) return(matrix(rest, nrow = 1))
    out <- list()
    for (v in vals) {
      if (v > rest + tol) next
      sub <- rec(k + 1, rest - v)
      if (is.matrix(sub)) out[[length(out)+1]] <- cbind(v, sub)
    }
    if (!length(out)) return(NULL)
    do.call(rbind, out)
  }
  ans <- rec(1, 1)
  ans[rowSums(ans) >= 1 - 1e-8 & rowSums(ans) <= 1 + 1e-8, , drop = FALSE]
}

.pick_threshold_constrained_local <- function(truth, prob, ths,
                                              min_spec, min_acc, target_sens){
  tab  <- .pr_table(truth, prob, ths)
  feas <- subset(tab, Specificity >= min_spec & Accuracy >= min_acc)

  if (nrow(feas) > 0) {
    feas90 <- subset(feas, Sensitivity >= target_sens)
    if (nrow(feas90) > 0) {
      J <- with(feas90, Sensitivity + Specificity - 1)
      best <- feas90[order(-J, -feas90$F1), ][1, , drop = FALSE]
      return(list(th = as.numeric(best$threshold), best_row = best, status = "HIT"))
    } else {
      J <- with(feas, Sensitivity + Specificity - 1)
      best <- feas[order(-feas$Sensitivity, -J, -feas$F1), ][1, , drop = FALSE]
      return(list(th = as.numeric(best$threshold), best_row = best, status = "LIMIT"))
    }
  }
  J <- with(tab, Sensitivity + Specificity - 1)
  best <- tab[order(-J, -tab$F1), ][1, , drop = FALSE]
  list(th = as.numeric(best$threshold), best_row = best, status = "NO_FEASIBLE")
}

build_prob_surrogate <- function(X_ref, p_ref) {
  if (!requireNamespace("glmnet", quietly = TRUE)) {
    stop("需要 glmnet 包：install.packages('glmnet')")
  }
  Xm  <- as.matrix(X_ref)
  fit <- glmnet::cv.glmnet(x = Xm, y = as.numeric(p_ref), family = "gaussian",
                           alpha = 0, nfolds = 5)
  function(newdata) {
    pr <- as.numeric(predict(fit, newx = as.matrix(newdata), s = "lambda.min"))
    pmin(pmax(pr, 1e-6), 1 - 1e-6)
  }
}


.cm_counts <- function(truth, pred){
  tp <- sum(truth=="Dyslexia" & pred=="Dyslexia", na.rm=TRUE)
  fp <- sum(truth=="Normal"   & pred=="Dyslexia", na.rm=TRUE)
  tn <- sum(truth=="Normal"   & pred=="Normal",   na.rm=TRUE)
  fn <- sum(truth=="Dyslexia" & pred=="Normal",   na.rm=TRUE)
  list(tp=tp, fp=fp, tn=tn, fn=fn)
}

.cm2matrix <- function(cnt){
  m <- matrix(c(cnt$tp, cnt$fp, cnt$fn, cnt$tn), nrow=2, byrow=TRUE)
  rownames(m) <- c("Pred=Dyslexia","Pred=Normal")
  colnames(m) <- c("Truth=Dyslexia","Truth=Normal")
  m
}

.metrics_from_counts <- function(cnt){
  n  <- cnt$tp + cnt$fp + cnt$tn + cnt$fn
  prec <- ifelse(cnt$tp+cnt$fp>0, cnt$tp/(cnt$tp+cnt$fp), NA_real_)
  rec  <- ifelse(cnt$tp+cnt$fn>0, cnt$tp/(cnt$tp+cnt$fn), NA_real_)
  spec <- ifelse(cnt$tn+cnt$fp>0, cnt$tn/(cnt$tn+cnt$fp), NA_real_)
  acc  <- (cnt$tp + cnt$tn) / n
  f1   <- ifelse(prec+rec>0, 2*prec*rec/(prec+rec), NA_real_)
  f2   <- ifelse(4*prec+rec>0, 5*prec*rec/(4*prec+rec), NA_real_)
  tibble::tibble(Accuracy=acc, Sensitivity=rec, Specificity=spec,
                 Precision=prec, F1=f1, F2=f2)
}
```



# define shap
```{r}

ensure_cols <- function(X_df, vec_needed){
  missing <- setdiff(vec_needed, colnames(X_df))
  if(length(missing)){
    for (v in missing) X_df[[v]] <- 0
  }
  X_df <- X_df[, vec_needed, drop = FALSE]
  return(as.matrix(X_df))
}

.prob_clip <- function(p, eps = 1e-6) pmax(pmin(p, 1 - eps), eps)
.logit <- function(p) qlogis(.prob_clip(p))
.sigmoid <- function(z) plogis(z)

compute_shap <- function(model, X_mat, y_factor,
                         pred_fun = NULL, pred_fun_prob = NULL,
                         top_features = NULL, top_n = 10,
                         nsim = 100, base_use = c("mean_logit","median_logit")) {
  base_use <- match.arg(base_use)
  f_prob <- if (!is.null(pred_fun_prob)) pred_fun_prob else
            if (!is.null(pred_fun))      pred_fun else
            stop("compute_shap() requires pred_fun_prob (or pred_fun) - a function returning the probability of 'Dyslexia'.")

  .prob_clip <- function(p, eps=1e-6) pmax(pmin(p, 1-eps), eps)
  .logit <- function(p) qlogis(.prob_clip(p))
  .sigmoid <- function(z) plogis(z)

  if (!is.null(top_features)) {
    X_mat <- X_mat[, intersect(top_features, colnames(X_mat)), drop=FALSE]
  }
  X_df <- as.data.frame(X_mat)

  pred_fun_logit <- function(m, newdata) .logit(f_prob(m, as.data.frame(newdata)))

  shap_logit <- fastshap::explain(
    object       = model,
    X            = X_df,
    pred_wrapper = pred_fun_logit,
    nsim         = nsim,
    adjust       = TRUE
  )
  if (is.null(shap_logit) || nrow(shap_logit) == 0) stop("SHAP matrix is empty.")

  base_logit_vec <- pred_fun_logit(model, X_df)
  base_logit <- if (base_use=="mean_logit") mean(base_logit_vec) else median(base_logit_vec)

  # logit-SHAP →  Δp
  delta_p <- apply(shap_logit, 2, function(sj) .sigmoid(base_logit + sj) - .sigmoid(base_logit))
  imp_df <- tibble::tibble(
    Feature       = colnames(delta_p),
    MeanAbsDeltaP = colMeans(abs(delta_p), na.rm = TRUE)
  ) |>
    dplyr::arrange(dplyr::desc(MeanAbsDeltaP)) |>
    dplyr::slice_head(n = top_n)

  shap_imp_plot <- ggplot2::ggplot(
      imp_df, ggplot2::aes(x = reorder(Feature, MeanAbsDeltaP), y = MeanAbsDeltaP * 100)
    ) +
    ggplot2::geom_col(width = 0.7, fill = "gray") +
    ggplot2::coord_flip() +
    ggplot2::labs(title = paste0("Top ", top_n, " SHAP (mean |Δp|)"),
                  x = "Feature", y = "Mean |Δp| (%)") +
    ggplot2::theme_minimal() +
    ggplot2::scale_y_continuous(limits = c(0, 50), breaks = seq(0, 50, by = 10),
                                expand = ggplot2::expansion(mult = c(0, 0.02)))

  sv_obj <- tryCatch(
  shapviz::shapviz(shap_logit, X = X_df, y = y_factor),
  error = function(e){
    message("[shapviz] Ignoring y:", e$message)
    shapviz::shapviz(shap_logit, X = X_df)  
    }
  )

  list(
    matrix   = delta_p, 
    sv       = sv_obj,
    plot     = shap_imp_plot,
    imp_df   = imp_df,
    base_logit = base_logit
  )
}
```



##  define LR
```{r}
run_lr_analysis <- function(train_data, test_data, dataset_name,
                            nfold = 10, smote_ratio = 0.6,
                            min_spec = 0.80,   
                            min_acc  = 0.80,    
                            target_sens = 0.90, 
                            th_grid = seq(0.01, 0.99, by = 0.01),
                            top_n_shap  = 10) {

  set.seed(111)

  if (is.factor(train_data$DD) && levels(train_data$DD)[1] != "Dyslexia")
    train_data$DD <- factor(train_data$DD, levels = c("Dyslexia","Normal"))
  if (is.factor(test_data$DD) && levels(test_data$DD)[1] != "Dyslexia")
    test_data$DD  <- factor(test_data$DD,  levels = c("Dyslexia","Normal"))

  for (nom in c("gender","Grade")) if (nom %in% names(train_data)) {
    train_data[[nom]] <- factor(as.character(train_data[[nom]]))
    test_data[[nom]]  <- factor(as.character(test_data[[nom]]), levels = levels(train_data[[nom]]))
  }

  # SMOTE
  rec_full  <- make_recipe(train_data, smote_ratio = smote_ratio)
  prep_full <- recipes::prep(rec_full, training = train_data)
  tr_full   <- recipes::juice(prep_full)                      # Training set (with SMOTE)
  te_full   <- recipes::bake(prep_full, new_data = test_data) # Test set (without SMOTE)
  y_tr      <- train_data$DD
  y_te      <- test_data$DD

  make_ConSens_summary <- function(spec_floor = 0.80,
                                   acc_floor  = 0.80,
                                   ths = seq(0.01, 0.99, by = 0.01)) {
  
   
    pr_table_local <- function(truth, prob, ths) {
      lev1 <- levels(truth)[1]  
      lev2 <- levels(truth)[2]
      out <- lapply(ths, function(t) {
        pred <- ifelse(prob >= t, lev1, lev2)
        pred <- factor(pred, levels = c(lev1, lev2))
  
        tp <- sum(truth==lev1 & pred==lev1)
        fp <- sum(truth==lev2 & pred==lev1)
        tn <- sum(truth==lev2 & pred==lev2)
        fn <- sum(truth==lev1 & pred==lev2)
  
        se  <- ifelse(tp+fn>0, tp/(tp+fn), 0)             
        sp  <- ifelse(tn+fp>0, tn/(tn+fp), 0)
        acc <- ifelse(tp+fp+tn+fn>0, (tp+tn)/(tp+fp+tn+fn), 0)
        data.frame(threshold = t, Sensitivity = se, Specificity = sp, Accuracy = acc)
      })
      do.call(rbind, out)
    }
    function(data, lev = NULL, model = NULL) {
      if (is.null(lev) || length(lev) != 2)    return(c(ConSens = 0))
      if (!lev[1] %in% names(data))            return(c(ConSens = 0))
      if (!("obs" %in% names(data)))           return(c(ConSens = 0))
  
      truth <- factor(data$obs, levels = lev)
      prob  <- suppressWarnings(as.numeric(data[[lev[1]]]))
      if (!any(is.finite(prob)))               return(c(ConSens = 0))
  
      tab <- pr_table_local(truth, prob, ths)
  
      ok <- with(tab, (Specificity >= spec_floor) & (Accuracy >= acc_floor) &
                      is.finite(Sensitivity) & is.finite(Specificity) & is.finite(Accuracy))
      if (!any(ok)) return(c(ConSens = 0))
  
      cons <- max(tab$Sensitivity[ok], na.rm = TRUE)
      if (!is.finite(cons)) cons <- 0
  
      c(ConSens = as.numeric(cons))
    }
  }

  customSummary <- make_ConSens_summary(spec_floor = min_spec,
                                      acc_floor  = min_acc,
                                      ths        = th_grid)

  train_with_id <- dplyr::mutate(train_data, .row_id = dplyr::row_number())
  folds <- rsample::vfold_cv(train_with_id, v = nfold, strata = "DD")
  oof_lr <- numeric(nrow(train_with_id))

  for (k in seq_along(folds$splits)) {
    dat_tr <- rsample::analysis(folds$splits[[k]])
    dat_va <- rsample::assessment(folds$splits[[k]])
    idx_va <- dat_va$.row_id

    for (nom in c("gender","Grade")) if (nom %in% names(dat_tr)) {
      dat_tr[[nom]] <- factor(as.character(dat_tr[[nom]]))
      dat_va[[nom]] <- factor(as.character(dat_va[[nom]]), levels = levels(dat_tr[[nom]]))
    }

    rec_k  <- make_recipe(dplyr::select(dat_tr, - .row_id), smote_ratio = smote_ratio)
    prep_k <- recipes::prep(rec_k, training = dat_tr)
    tr_k   <- recipes::juice(prep_k)                               
    va_k   <- recipes::bake(prep_k, new_data = dplyr::select(dat_va, - .row_id)) 

    ctrl <- caret::trainControl(method = "cv", number = 5,
                                classProbs = TRUE,
                                summaryFunction = customSummary,
                                verboseIter = FALSE)
    grid <- expand.grid(alpha = c(0, 0.5, 1),
                        lambda = 10^seq(-4, 0, length.out = 40))
    
    fit_lr <- caret::train(DD ~ ., data = tr_k,
                           method = "glmnet",
                           trControl = ctrl,
                           tuneGrid = grid,
                           metric = "ConSens")  

    oof_pred <- as.numeric(predict(fit_lr, va_k, type = "prob")[, "Dyslexia"])
    oof_lr[idx_va] <- oof_pred
  }
  
  
  pick_threshold_constrained <- function(truth, prob,
                                       ths = th_grid,
                                       target_sens_val = target_sens,
                                       min_spec_val    = min_spec,
                                       min_acc_val     = min_acc) {
  tab  <- .pr_table(truth, prob, ths)
  feas <- subset(tab, Specificity >= min_spec_val & Accuracy >= min_acc_val)

  feas_target <- subset(feas, Sensitivity >= target_sens_val)
  if (nrow(feas_target)) {
    J <- with(feas_target, Sensitivity + Specificity - 1)
    best <- feas_target[order(-J, -feas_target$F1), ][1, , drop = FALSE]
    return(list(th = best$threshold, table = tab, best_row = best, met_all = TRUE))
  }

  if (nrow(feas)) {
    J <- with(feas, Sensitivity + Specificity - 1)
    best <- feas[order(-feas$Sensitivity, -J, -feas$F1), ][1, , drop = FALSE]
    return(list(th = best$threshold, table = tab, best_row = best, met_all = FALSE))
  }

  J <- with(tab, Sensitivity + Specificity - 1)
  best <- tab[order(-J, -tab$F1), ][1, , drop = FALSE]
  warning(sprintf("Threshold satisfying Sp>=%.2f & Acc>=%.2f not found; returning max J point (Se=%.3f, Sp=%.3f, Acc=%.3f, thr=%.2f).",
                  min_spec_val, min_acc_val, best$Sensitivity, best$Specificity, best$Accuracy, best$threshold))
  list(th = best$threshold, table = tab, best_row = best, met_all = FALSE)
}

  thr_pick <- pick_threshold_constrained(
    train_data$DD, oof_lr,
    ths = th_grid,
    target_sens_val = target_sens,
    min_spec_val    = min_spec,
    min_acc_val     = min_acc
  )
  th_star <- as.numeric(thr_pick$th) 
  
  sel_idx <- which.min(abs(thr_pick$table$threshold - th_star))
  thresh_selected_row <- thr_pick$table[sel_idx, , drop = FALSE]
  

  ctrl_all <- caret::trainControl(method = "cv", number = 10,
                                  classProbs = TRUE,
                                  summaryFunction = customSummary)
  grid_all <- expand.grid(alpha = c(0, 0.5, 1),
                          lambda = 10^seq(-4, 0, length.out = 60))
  
  fit_lr_full <- caret::train(DD ~ ., data = tr_full,
                              method = "glmnet",
                              trControl = ctrl_all,
                              tuneGrid = grid_all,
                              metric = "ConSens")

  p_lr_test <- as.numeric(predict(fit_lr_full, te_full, type = "prob")[, "Dyslexia"])
  truth_te  <- factor(y_te, levels = c("Dyslexia","Normal"))
  pred_te   <- factor(ifelse(p_lr_test >= th_star, "Dyslexia","Normal"),
                      levels = c("Dyslexia","Normal"))


  tp <- sum(truth_te=="Dyslexia" & pred_te=="Dyslexia")
  fp <- sum(truth_te=="Normal"   & pred_te=="Dyslexia")
  tn <- sum(truth_te=="Normal"   & pred_te=="Normal")
  fn <- sum(truth_te=="Dyslexia" & pred_te=="Normal")
  Acc <- tibble::tibble(
    Accuracy    = (tp+tn)/max(1,tp+fp+tn+fn),
    Sensitivity = ifelse(tp+fn>0, tp/(tp+fn), NA_real_),
    Specificity = ifelse(tn+fp>0, tn/(tn+fp), NA_real_),
    Precision   = ifelse(tp+fp>0, tp/(tp+fp), NA_real_),
    F1          = ifelse((2*tp+fp+fn)>0, 2*tp/(2*tp+fp+fn), NA_real_),
    F2          = ifelse((5*tp+4*fn+fp)>0, 5*tp/(5*tp+4*fn+fp), NA_real_)
  )

  conf_df <- tibble::tibble(truth = truth_te, pred = pred_te)
  conf_mx <- yardstick::conf_mat(conf_df, truth = truth, estimate = pred)

  metrics <- tibble::tibble(
    Dataset         = dataset_name,
    Threshold       = round(th_star, 2),       
    Threshold_exact = signif(th_star, 12),    
    AUC         = as.numeric(yardstick::roc_auc_vec(truth_te, p_lr_test, event_level = "first")),
    Accuracy    = Acc$Accuracy,
    Sensitivity = Acc$Sensitivity,
    Specificity = Acc$Specificity,
    Precision   = Acc$Precision,
    F1          = Acc$F1,
    F2          = Acc$F2
  )

  # SHAP
  X_for_shap <- tr_full %>% dplyr::select(-DD) %>% as.data.frame()
  pred_fun_lr_prob <- function(model, newdata){
    newdata <- as.data.frame(newdata)
    as.numeric(predict(model, newdata = newdata, type = "prob")[, "Dyslexia"])
  }
  lr_shap <- compute_shap(
    model        = fit_lr_full,
    X_mat        = X_for_shap,
    y_factor     = y_tr,
    pred_fun_prob= pred_fun_lr_prob,
    nsim         = 100,
    top_n        = top_n_shap
  )

  
  best_oof_row <- as.data.frame(thr_pick$best_row)
  thr_rep      <- as.numeric(best_oof_row$threshold)

  truth <- factor(truth_te, levels = c("Dyslexia","Normal"))
  pred  <- factor(ifelse(p_lr_test >= thr_rep, "Dyslexia","Normal"),
                  levels = c("Dyslexia","Normal"))

  tp <- sum(truth=="Dyslexia" & pred=="Dyslexia")
  fp <- sum(truth=="Normal"   & pred=="Dyslexia")
  tn <- sum(truth=="Normal"   & pred=="Normal")
  fn <- sum(truth=="Dyslexia" & pred=="Normal")

  rep_metrics_test <- data.frame(
    Thr_from_OOF = thr_rep,
    OOF_Sens = best_oof_row$Sensitivity,
    OOF_Spec = best_oof_row$Specificity,
    OOF_Acc  = best_oof_row$Accuracy,
    OOF_F1   = best_oof_row$F1,
    Test_Sens = ifelse(tp+fn>0, tp/(tp+fn), NA_real_),
    Test_Spec = ifelse(tn+fp>0, tn/(tn+fp), NA_real_),
    Test_Acc  = (tp+tn)/max(1, tp+fp+tn+fn),
    Test_Prec = ifelse(tp+fp>0, tp/(tp+fp), NA_real_),
    Test_F1   = ifelse((2*tp+fp+fn)>0, 2*tp/(2*tp+fp+fn), NA_real_)
  )

  pick_rep <- list(row = best_oof_row)

  list(
    metrics      = metrics,
    conf_mat     = conf_mx,
    probs_test   = p_lr_test,
    truth_test   = truth_te,
    threshold    = th_star,
    thresh_table = thr_pick$table,     
    thresh_best  = thr_pick$best_row,    
    report_best_row     = pick_rep$row,  
    report_test_at_best = rep_metrics_test,
    threshold_exact = th_star,
    thresh_selected_row = thresh_selected_row,
    params = list(min_spec = min_spec, min_acc = min_acc,
                target_sens = target_sens, th_grid = th_grid),
    roc_plot     = {
      roc_obj <- pROC::roc(response = truth_te, predictor = p_lr_test,
                           levels = c("Dyslexia","Normal"), direction = ">")
      pROC::ggroc(roc_obj) +
        ggplot2::geom_abline(linetype = "dashed", colour = "grey") +
        ggplot2::annotate("text", x = 0.7, y = 0.3,
                          label = paste("AUC =", round(pROC::auc(roc_obj), 3))) +
        ggplot2::labs(title = paste("ROC Curve - LR -", dataset_name),
                      x = "False Positive Rate", y = "True Positive Rate") +
        ggplot2::theme_bw()
    },
    shap_plot    = lr_shap$plot,
    shap_values  = lr_shap$matrix,
    shap_imp_df  = lr_shap$imp_df,
    model        = fit_lr_full,
    x_test       = te_full %>% dplyr::select(-DD)
  )

}

```


##  define RF
```{r}
run_rf_analysis <- function(train_data, test_data, dataset_name,
                            nfold = 10, smote_ratio = 0.6,
                            min_spec = 0.80, min_acc = 0.80, target_sens = 0.90,
                            th_grid = seq(0.01,0.99,0.01),
                            num.trees = 500,
                            top_n_shap = 10){
  set.seed(111)
  for (nm in c("DD","gender","Grade")) if (nm %in% names(train_data)){
    if (nm=="DD"){
      train_data$DD <- factor(train_data$DD, levels=c("Dyslexia","Normal"))
      test_data$DD  <- factor(test_data$DD,  levels=c("Dyslexia","Normal"))
    } else {
      train_data[[nm]] <- factor(as.character(train_data[[nm]]))
      test_data[[nm]]  <- factor(as.character(test_data[[nm]]), levels=levels(train_data[[nm]]))
    }
  }

  rec_full  <- make_recipe(train_data, smote_ratio = smote_ratio)
  prep_full <- recipes::prep(rec_full, training=train_data)
  tr_full   <- recipes::juice(prep_full)
  te_full   <- recipes::bake(prep_full, new_data=test_data)
  y_tr <- train_data$DD; y_te <- test_data$DD


  customSummary <- make_ConSens_summary(min_spec, min_acc, th_grid)

  train_with_id <- dplyr::mutate(train_data, .row_id = dplyr::row_number())
  folds <- rsample::vfold_cv(train_with_id, v = nfold, strata = "DD")
  oof <- numeric(nrow(train_with_id))

  for (k in seq_along(folds$splits)){
    dat_tr <- rsample::analysis(folds$splits[[k]])
    dat_va <- rsample::assessment(folds$splits[[k]])
    idx_va <- dat_va$.row_id

    for (nm in c("gender","Grade")) if (nm %in% names(dat_tr)){
      dat_tr[[nm]] <- factor(as.character(dat_tr[[nm]]))
      dat_va[[nm]] <- factor(as.character(dat_va[[nm]]), levels=levels(dat_tr[[nm]]))
    }

    rec_k  <- make_recipe(dplyr::select(dat_tr, - .row_id), smote_ratio=smote_ratio)
    prep_k <- recipes::prep(rec_k, training=dat_tr)
    tr_k   <- recipes::juice(prep_k)
    va_k   <- recipes::bake(prep_k, new_data=dplyr::select(dat_va, - .row_id))

    ctrl <- caret::trainControl(method="cv", number=5, classProbs=TRUE, summaryFunction=customSummary)
    grid <- expand.grid(
      mtry          = pmax(1, round(c(0.1, 0.3, 0.5) * (ncol(tr_k)-1))), 
      splitrule     = c("gini","extratrees"),
      min.node.size = c(1, 5, 10)
    )

    fit <- caret::train(DD ~ ., data=tr_k,
                        method="ranger", trControl=ctrl, tuneGrid=grid,
                        metric="ConSens",
                        num.trees=num.trees,  importance="impurity")
    oof[idx_va] <- as.numeric(predict(fit, va_k, type="prob")[,"Dyslexia"])
  }

  thr_pick <- pick_threshold_constrained(train_data$DD, oof, th_grid, target_sens, min_spec, min_acc)
  th_star  <- as.numeric(thr_pick$th)
  sel_idx  <- which.min(abs(thr_pick$table$threshold - th_star))
  thresh_selected_row <- thr_pick$table[sel_idx, , drop = FALSE]

  ctrl_all <- caret::trainControl(method="cv", number=10, classProbs=TRUE, summaryFunction=customSummary)
  grid_all <- expand.grid(
    mtry          = pmax(1, round(c(0.1, 0.3, 0.5) * (ncol(tr_full)-1))),
    splitrule     = c("gini","extratrees"),
    min.node.size = c(1, 5, 10)
  )
  fit_full <- caret::train(DD ~ ., data=tr_full, method="ranger",
                           trControl=ctrl_all, tuneGrid=grid_all, metric="ConSens",
                           num.trees=num.trees, importance="impurity")

  p_test <- as.numeric(predict(fit_full, te_full, type="prob")[,"Dyslexia"])
  truth  <- factor(y_te, levels=c("Dyslexia","Normal"))
  pred   <- factor(ifelse(p_test>=th_star, "Dyslexia","Normal"), levels=c("Dyslexia","Normal"))

  tp <- sum(truth=="Dyslexia" & pred=="Dyslexia"); fp <- sum(truth=="Normal" & pred=="Dyslexia")
  tn <- sum(truth=="Normal" & pred=="Normal");     fn <- sum(truth=="Dyslexia" & pred=="Normal")

  metrics <- tibble::tibble(
    Dataset         = dataset_name,
    Threshold       = round(th_star,2),
    Threshold_exact = signif(th_star,12),
    AUC         = as.numeric(yardstick::roc_auc_vec(truth, p_test, event_level="first")),
    Accuracy    = (tp+tn)/(tp+fp+tn+fn),
    Sensitivity = ifelse(tp+fn>0, tp/(tp+fn), NA_real_),
    Specificity = ifelse(tn+fp>0, tn/(tn+fp), NA_real_),
    Precision   = ifelse(tp+fp>0, tp/(tp+fp), NA_real_),
    F1          = ifelse((2*tp+fp+fn)>0, 2*tp/(2*tp+fp+fn), NA_real_),
    F2          = ifelse((5*tp+4*fn+fp)>0, 5*tp/(5*tp+4*fn+fp), NA_real_)
  )

  conf_mx <- yardstick::conf_mat(tibble::tibble(truth=truth, pred=pred),
                                 truth=truth, estimate=pred)

  X_for_shap <- tr_full %>% dplyr::select(-DD) %>% as.data.frame()
  pred_fun_prob <- function(model, newdata){
    as.numeric(predict(model, newdata=as.data.frame(newdata), type="prob")[,"Dyslexia"])
  }
  rf_shap <- compute_shap(fit_full, X_for_shap, y_tr, pred_fun_prob, nsim=100, top_n=top_n_shap)

  list(
    metrics = metrics,
    conf_mat = conf_mx,
    probs_test = p_test,
    truth_test = truth,
    threshold = th_star,
    thresh_table = thr_pick$table,
    thresh_best = thr_pick$best_row,
    thresh_selected_row = thresh_selected_row,
    threshold_exact = th_star,
    params = list(min_spec=min_spec, min_acc=min_acc, target_sens=target_sens, th_grid=th_grid),
    shap_plot = rf_shap$plot,
    shap_values_logit = rf_shap$shap_logit,
    shap_values_deltap = rf_shap$shap_deltap,
    shap_imp_deltap = rf_shap$imp_deltap,
    model = fit_full,
    x_test = te_full %>% dplyr::select(-DD)
  )
}



```



#  define XGBoost
```{r}
run_xgb_analysis <- function(train_data, test_data, dataset_name,
                             nfold = 10, smote_ratio = 0.6,
                             min_spec = 0.80, min_acc = 0.80, target_sens = 0.90,
                             th_grid = seq(0.01,0.99,0.01),
                             top_n_shap = 10){
  set.seed(111)
  for (nm in c("DD","gender","Grade")) if (nm %in% names(train_data)){
    if (nm=="DD"){
      train_data$DD <- factor(train_data$DD, levels=c("Dyslexia","Normal"))
      test_data$DD  <- factor(test_data$DD,  levels=c("Dyslexia","Normal"))
    } else {
      train_data[[nm]] <- factor(as.character(train_data[[nm]]))
      test_data[[nm]]  <- factor(as.character(test_data[[nm]]), levels=levels(train_data[[nm]]))
    }
  }

  rec_full  <- make_recipe(train_data, smote_ratio = smote_ratio)
  prep_full <- recipes::prep(rec_full, training=train_data)
  tr_full   <- recipes::juice(prep_full)
  te_full   <- recipes::bake(prep_full, new_data=test_data)
  y_tr <- train_data$DD; y_te <- test_data$DD

  customSummary <- make_ConSens_summary(min_spec, min_acc, th_grid)

  train_with_id <- dplyr::mutate(train_data, .row_id = dplyr::row_number())
  folds <- rsample::vfold_cv(train_with_id, v = nfold, strata = "DD")
  oof <- numeric(nrow(train_with_id))

  for (k in seq_along(folds$splits)){
    dat_tr <- rsample::analysis(folds$splits[[k]])
    dat_va <- rsample::assessment(folds$splits[[k]])
    idx_va <- dat_va$.row_id

    for (nm in c("gender","Grade")) if (nm %in% names(dat_tr)){
      dat_tr[[nm]] <- factor(as.character(dat_tr[[nm]]))
      dat_va[[nm]] <- factor(as.character(dat_va[[nm]]), levels=levels(dat_tr[[nm]]))
    }

    rec_k  <- make_recipe(dplyr::select(dat_tr, - .row_id), smote_ratio=smote_ratio)
    prep_k <- recipes::prep(rec_k, training=dat_tr)
    tr_k   <- recipes::juice(prep_k)
    va_k   <- recipes::bake(prep_k, new_data=dplyr::select(dat_va, - .row_id))

    ctrl <- caret::trainControl(method="cv", number=5, classProbs=TRUE, summaryFunction=customSummary)
    grid <- expand.grid(
      nrounds            = c(200, 400),
      max_depth          = c(3, 5, 7),
      eta                = c(0.05, 0.1, 0.2),
      gamma              = c(0, 1),
      colsample_bytree   = c(0.7, 1.0),
      min_child_weight   = c(1, 5),
      subsample          = c(0.7, 1.0)
    )
    fit <- caret::train(DD ~ ., data=tr_k,
                        method="xgbTree", trControl=ctrl, tuneGrid=grid,
                        metric="ConSens", verbosity=0)
    oof[idx_va] <- as.numeric(predict(fit, va_k, type="prob")[,"Dyslexia"])
  }

  thr_pick <- pick_threshold_constrained(train_data$DD, oof, th_grid, target_sens, min_spec, min_acc)
  th_star  <- as.numeric(thr_pick$th)
  sel_idx  <- which.min(abs(thr_pick$table$threshold - th_star))
  thresh_selected_row <- thr_pick$table[sel_idx, , drop = FALSE]


  ctrl_all <- caret::trainControl(method="cv", number=10, classProbs=TRUE, summaryFunction=customSummary)
  grid_all <- grid  
  fit_full <- caret::train(DD ~ ., data=tr_full,
                           method="xgbTree", trControl=ctrl_all, tuneGrid=grid_all,
                           metric="ConSens", verbosity=0)

  p_test <- as.numeric(predict(fit_full, te_full, type="prob")[,"Dyslexia"])
  truth  <- factor(y_te, levels=c("Dyslexia","Normal"))
  pred   <- factor(ifelse(p_test>=th_star, "Dyslexia","Normal"), levels=c("Dyslexia","Normal"))

  tp <- sum(truth=="Dyslexia" & pred=="Dyslexia"); fp <- sum(truth=="Normal" & pred=="Dyslexia")
  tn <- sum(truth=="Normal" & pred=="Normal");     fn <- sum(truth=="Dyslexia" & pred=="Normal")

  metrics <- tibble::tibble(
    Dataset         = dataset_name,
    Threshold       = round(th_star,2),
    Threshold_exact = signif(th_star,12),
    AUC         = as.numeric(yardstick::roc_auc_vec(truth, p_test, event_level="first")),
    Accuracy    = (tp+tn)/(tp+fp+tn+fn),
    Sensitivity = ifelse(tp+fn>0, tp/(tp+fn), NA_real_),
    Specificity = ifelse(tn+fp>0, tn/(tn+fp), NA_real_),
    Precision   = ifelse(tp+fp>0, tp/(tp+fp), NA_real_),
    F1          = ifelse((2*tp+fp+fn)>0, 2*tp/(2*tp+fp+fn), NA_real_),
    F2          = ifelse((5*tp+4*fn+fp)>0, 5*tp/(5*tp+4*fn+fp), NA_real_)
  )
  conf_mx <- yardstick::conf_mat(tibble::tibble(truth=truth, pred=pred),
                                 truth=truth, estimate=pred)


  X_for_shap <- tr_full %>% dplyr::select(-DD) %>% as.data.frame()
  pred_fun_prob <- function(model, newdata){
    as.numeric(predict(model, newdata=as.data.frame(newdata), type="prob")[,"Dyslexia"])
  }
  xgb_shap <- compute_shap(fit_full, X_for_shap, y_tr, pred_fun_prob, nsim=100, top_n=top_n_shap)

  list(
    metrics=metrics, conf_mat=conf_mx, probs_test=p_test, truth_test=truth,
    threshold=th_star, thresh_table=thr_pick$table, thresh_best=thr_pick$best_row,
    thresh_selected_row = thresh_selected_row,
    threshold_exact=th_star,
    params=list(min_spec=min_spec, min_acc=min_acc, target_sens=target_sens, th_grid=th_grid),
    shap_plot=xgb_shap$plot,
    shap_values_logit=xgb_shap$shap_logit,
    shap_values_deltap=xgb_shap$shap_deltap,
    shap_imp_deltap=xgb_shap$imp_deltap,
    model=fit_full, x_test=te_full %>% dplyr::select(-DD)
  )
}


```


##  define MLP
```{r}
run_mlp_analysis <- function(train_data, test_data, dataset_name,
                             nfold = 10, smote_ratio = 0.6,
                             min_spec = 0.80,   
                             min_acc  = 0.80,    
                             target_sens = 0.90, 
                             th_grid = seq(0.01, 0.99, by = 0.01),
                             epochs = 15, batch_size = 32,
                             units1 = c(64, 128),
                             units2 = c(32, 64),
                             l2_reg = c(1e-4, 1e-5),
                             dropout = c(0.2),
                             lr = c(1e-3),
                             top_n_shap = 10) {

  set.seed(111)
  if (requireNamespace("tensorflow", quietly = TRUE)) {
    try({ tensorflow::tf$random$set_seed(111) }, silent = TRUE)
  }

  to_pos_first <- function(x){
    if (is.factor(x) && levels(x)[1] != "Dyslexia")
      factor(x, levels = c("Dyslexia","Normal")) else x
  }
  train_data$DD <- to_pos_first(train_data$DD)
  test_data$DD  <- to_pos_first(test_data$DD)

  for (nom in c("gender","Grade")) if (nom %in% names(train_data)) {
    train_data[[nom]] <- factor(as.character(train_data[[nom]]))
    test_data[[nom]]  <- factor(as.character(test_data[[nom]]),
                                levels = levels(train_data[[nom]]))
  }

  rec_full  <- make_recipe(train_data, smote_ratio = smote_ratio)
  prep_full <- recipes::prep(rec_full, training = train_data)
  tr_full   <- recipes::juice(prep_full)                      
  te_full   <- recipes::bake(prep_full, new_data = test_data) 
  y_tr      <- train_data$DD
  y_te      <- test_data$DD

  pick_threshold_constrained <- function(truth, prob,
                                         ths = th_grid,
                                         target_sens_val = target_sens,
                                         min_spec_val    = min_spec,
                                         min_acc_val     = min_acc) {
    tab  <- .pr_table(truth, prob, ths)
    feas <- subset(tab, Specificity >= min_spec_val & Accuracy >= min_acc_val)

    feas_target <- subset(feas, Sensitivity >= target_sens_val)
    if (nrow(feas_target)) {
      J <- with(feas_target, Sensitivity + Specificity - 1)
      best <- feas_target[order(-J, -feas_target$F1), ][1, , drop = FALSE]
      return(list(th = best$threshold, table = tab, best_row = best, met_all = TRUE))
    }

    if (nrow(feas)) {
      J <- with(feas, Sensitivity + Specificity - 1)
      best <- feas[order(-feas$Sensitivity, -J, -feas$F1), ][1, , drop = FALSE]
      return(list(th = best$threshold, table = tab, best_row = best, met_all = FALSE))
    }

    J <- with(tab, Sensitivity + Specificity - 1)
    best <- tab[order(-J, -tab$F1), ][1, , drop = FALSE]
    warning(sprintf("Threshold satisfying Sp>=%.2f & Acc>=%.2f not found; returning max J point (Se=%.3f, Sp=%.3f, Acc=%.3f, thr=%.2f).",
                    min_spec_val, min_acc_val, best$Sensitivity, best$Specificity, best$Accuracy, best$threshold))
    list(th = best$threshold, table = tab, best_row = best, met_all = FALSE)
  }

  build_mlp <- function(input_dim, u1, u2, dp, l2v, lr){
    k <- keras::backend()
    keras::keras_model_sequential() |>
      keras::layer_dense(units = u1, activation = "relu",
                         input_shape = input_dim,
                         kernel_regularizer = keras::regularizer_l2(l2v)) |>
      keras::layer_dropout(rate = dp) |>
      keras::layer_dense(units = u2, activation = "relu",
                         kernel_regularizer = keras::regularizer_l2(l2v)) |>
      keras::layer_dropout(rate = dp) |>
      keras::layer_dense(units = 1, activation = "sigmoid") |>
      keras::compile(
        optimizer = keras::optimizer_rmsprop(learning_rate = lr),
        loss = "binary_crossentropy"
      )
  }

  y01 <- function(f) as.numeric(f == "Dyslexia")

  grid <- expand.grid(
    u1 = units1, u2 = units2,
    dp = dropout, l2v = l2_reg, lrv = lr,
    KEEP.OUT.ATTRS = FALSE
  )

  train_with_id <- dplyr::mutate(train_data, .row_id = dplyr::row_number())
  folds <- rsample::vfold_cv(train_with_id, v = nfold, strata = "DD")
  oof_prob <- numeric(nrow(train_with_id))
  winners  <- list()  

  score_ConSens <- function(truth, prob){
    tab <- .pr_table(truth, prob, ths = th_grid)
    ok  <- with(tab, Specificity >= min_spec & Accuracy >= min_acc)
    if (!any(ok)) return(0)
    max(tab$Sensitivity[ok], na.rm = TRUE)
  }

  for (k in seq_along(folds$splits)) {
    dat_tr <- rsample::analysis(folds$splits[[k]])
    dat_va <- rsample::assessment(folds$splits[[k]])
    idx_va <- dat_va$.row_id

    rec_k  <- make_recipe(dplyr::select(dat_tr, - .row_id), smote_ratio = smote_ratio)
    prep_k <- recipes::prep(rec_k, training = dat_tr)
    tr_k   <- recipes::juice(prep_k)                                     
    va_k   <- recipes::bake(prep_k, new_data = dplyr::select(dat_va, - .row_id))

    X_tr <- as.matrix(tr_k |> dplyr::select(-DD))
    z_tr <- y01(tr_k$DD)
    X_va <- as.matrix(va_k |> dplyr::select(-DD))
    z_va <- factor(va_k$DD, levels = c("Dyslexia","Normal"))

    grid_scores <- c()
    preds_hold  <- list()
    for (gi in seq_len(nrow(grid))) {
      g <- grid[gi, ]
      mdl <- build_mlp(ncol(X_tr), g$u1, g$u2, g$dp, g$l2v, g$lrv)
      history <- mdl %>% keras::fit(
        x = X_tr, y = z_tr,
        epochs = epochs, batch_size = batch_size,
        validation_split = 0, verbose = 0
      )
      p_va <- as.numeric(mdl %>% predict(X_va, verbose = 0))
      preds_hold[[gi]] <- p_va
      grid_scores[gi]  <- score_ConSens(z_va, p_va)
      try({ keras::k_clear_session() }, silent = TRUE)
    }
    best_idx <- which.max(grid_scores)
    oof_prob[idx_va] <- preds_hold[[best_idx]]
    winners[[k]] <- dplyr::bind_cols(grid[best_idx, ], tibble::tibble(ConSens = max(grid_scores, na.rm = TRUE)))
  }

  win_tab <- dplyr::bind_rows(winners) |>
    dplyr::group_by(u1, u2, dp, l2v, lrv) |>
    dplyr::summarise(n = dplyr::n(), meanConSens = mean(ConSens), .groups = "drop") |>
    dplyr::arrange(dplyr::desc(n), dplyr::desc(meanConSens))
  best_cfg <- win_tab[1, ]

  thr_pick <- pick_threshold_constrained(
    truth = train_data$DD, prob = oof_prob,
    ths = th_grid, target_sens_val = target_sens,
    min_spec_val = min_spec, min_acc_val = min_acc
  )
  th_star <- as.numeric(thr_pick$th)
  sel_idx <- which.min(abs(thr_pick$table$threshold - th_star))
  thresh_selected_row <- thr_pick$table[sel_idx, , drop = FALSE]

  X_tr_full <- as.matrix(tr_full |> dplyr::select(-DD))
  z_tr_full <- y01(tr_full$DD)
  X_te      <- as.matrix(te_full |> dplyr::select(-DD))
  truth_te  <- factor(y_te, levels = c("Dyslexia","Normal"))

  mdl_full <- build_mlp(
    input_dim = ncol(X_tr_full),
    u1 = best_cfg$u1, u2 = best_cfg$u2,
    dp = best_cfg$dp, l2v = best_cfg$l2v, lr = best_cfg$lrv
  )
  history <- mdl_full %>% keras::fit(
    x = X_tr_full, y = z_tr_full,
    epochs = epochs, batch_size = batch_size,
    validation_split = 0, verbose = 0
  )
  p_test <- as.numeric(mdl_full %>% predict(X_te, verbose = 0))
  pred_te <- factor(ifelse(p_test >= th_star, "Dyslexia","Normal"),
                    levels = c("Dyslexia","Normal"))

  tp <- sum(truth_te=="Dyslexia" & pred_te=="Dyslexia")
  fp <- sum(truth_te=="Normal"   & pred_te=="Dyslexia")
  tn <- sum(truth_te=="Normal"   & pred_te=="Normal")
  fn <- sum(truth_te=="Dyslexia" & pred_te=="Normal")
  Acc <- tibble::tibble(
    Accuracy    = (tp+tn)/max(1,tp+fp+tn+fn),
    Sensitivity = ifelse(tp+fn>0, tp/(tp+fn), NA_real_),
    Specificity = ifelse(tn+fp>0, tn/(tn+fp), NA_real_),
    Precision   = ifelse(tp+fp>0, tp/(tp+fp), NA_real_),
    F1          = ifelse((2*tp+fp+fn)>0, 2*tp/(2*tp+fp+fn), NA_real_),
    F2          = ifelse((5*tp+4*fn+fp)>0, 5*tp/(5*tp+4*fn+fp), NA_real_)
  )
  conf_df <- tibble::tibble(truth = truth_te, pred = pred_te)
  conf_mx <- yardstick::conf_mat(conf_df, truth = truth, estimate = pred)

  metrics <- tibble::tibble(
    Dataset         = dataset_name,
    Threshold       = round(th_star, 2),
    Threshold_exact = signif(th_star, 12),
    AUC         = as.numeric(yardstick::roc_auc_vec(truth_te, p_test, event_level = "first")),
    Accuracy    = Acc$Accuracy,
    Sensitivity = Acc$Sensitivity,
    Specificity = Acc$Specificity,
    Precision   = Acc$Precision,
    F1          = Acc$F1,
    F2          = Acc$F2
  )

  pred_fun_mlp <- function(model, newdata){
    as.numeric(model %>% predict(as.matrix(newdata), verbose = 0))
  }
  shap_out <- compute_shap(
    model         = mdl_full,
    X_mat         = tr_full |> dplyr::select(-DD) |> as.data.frame(),
    y_factor      = y_tr,
    pred_fun_prob = pred_fun_mlp,
    nsim          = 50,
    top_n         = top_n_shap
  )

  best_oof_row <- as.data.frame(thr_pick$best_row)
  thr_rep      <- as.numeric(best_oof_row$threshold)
  pred_rep     <- factor(ifelse(p_test >= thr_rep, "Dyslexia","Normal"),
                         levels = c("Dyslexia","Normal"))
  tp2 <- sum(truth_te=="Dyslexia" & pred_rep=="Dyslexia")
  fp2 <- sum(truth_te=="Normal"   & pred_rep=="Dyslexia")
  tn2 <- sum(truth_te=="Normal"   & pred_rep=="Normal")
  fn2 <- sum(truth_te=="Dyslexia" & pred_rep=="Normal")
  rep_metrics_test <- data.frame(
    Thr_from_OOF = thr_rep,
    OOF_Sens = best_oof_row$Sensitivity,
    OOF_Spec = best_oof_row$Specificity,
    OOF_Acc  = best_oof_row$Accuracy,
    OOF_F1   = best_oof_row$F1,
    Test_Sens = ifelse(tp2+fn2>0, tp2/(tp2+fn2), NA_real_),
    Test_Spec = ifelse(tn2+fp2>0, tn2/(tn2+fp2), NA_real_),
    Test_Acc  = (tp2+tn2)/max(1, tp2+fp2+tn2+fn2),
    Test_Prec = ifelse(tp2+fp2>0, tp2/(tp2+fp2), NA_real_),
    Test_F1   = ifelse((2*tp2+fp2+fn2)>0, 2*tp2/(2*tp2+fp2+fn2), NA_real_)
  )

  list(
    metrics      = metrics,
    conf_mat     = conf_mx,
    probs_test   = p_test,
    truth_test   = truth_te,
    threshold    = th_star,
    thresh_table = thr_pick$table,
    thresh_best  = thr_pick$best_row,
    report_best_row     = best_oof_row,
    report_test_at_best = rep_metrics_test,
    threshold_exact     = th_star,
    thresh_selected_row = thresh_selected_row,
    params = list(min_spec = min_spec, min_acc = min_acc,
                  target_sens = target_sens, th_grid = th_grid,
                  epochs = epochs, batch_size = batch_size,
                  best_cfg = best_cfg, win_table = win_tab),
    roc_plot     = {
      roc_obj <- pROC::roc(response = truth_te, predictor = p_test,
                           levels = c("Dyslexia","Normal"), direction = ">")
      pROC::ggroc(roc_obj) +
        ggplot2::geom_abline(linetype = "dashed", colour = "grey50") +
        ggplot2::annotate("text", x = 0.7, y = 0.3,
                          label = paste("AUC =", round(pROC::auc(roc_obj), 3))) +
        ggplot2::labs(title = paste("ROC Curve - MLP -", dataset_name),
                      x = "False Positive Rate", y = "True Positive Rate") +
        ggplot2::theme_bw()
    },
    shap_plot    = shap_out$plot,         
    shap_values  = shap_out$shap_deltap,  
    shap_imp_df  = shap_out$imp_deltap,    
    model        = mdl_full,
    x_test       = te_full |> dplyr::select(-DD)
  )
}

```

# define vote
```{r}
vote_optimize <- function(i,
                          step_w = 0.10,
                          th_grid = seq(0.01, 0.99, 0.01),
                          min_spec = 0.80, min_acc = 0.80, target_sens = 0.90,
                          top_n_shap = 10,
                          verbose = TRUE){

  lr  <- get_result_obj("LR",  i)
  rf  <- get_result_obj("RF",  i)
  xgb <- get_result_obj("XGB", i)
  mlp <- get_result_obj("MLP", i)
  truth <- lr$truth_test %||% rf$truth_test %||% xgb$truth_test %||% mlp$truth_test
  if (is.null(truth)) stop(sprintf("Dataset %d: truth_test not found.", i))
  truth <- factor(truth, levels = c("Dyslexia","Normal"))

  X_ref <- lr$x_test %||% rf$x_test %||% xgb$x_test %||% mlp$x_test
  X_ref <- as.data.frame(X_ref)

  parts <- list()
  if (!is.null(lr))  { p <- fetch_probs_safe(lr);  if (!is.null(p)) parts$LR  <- p }
  if (!is.null(rf))  { p <- fetch_probs_safe(rf);  if (!is.null(p)) parts$RF  <- p }
  if (!is.null(xgb)) { p <- fetch_probs_safe(xgb); if (!is.null(p)) parts$XGB <- p }
  if (!is.null(mlp)) { p <- fetch_probs_safe(mlp); if (!is.null(p)) parts$MLP <- p }
  if (!length(parts)) stop(sprintf("Dataset %d: No available single-model probabilities (probs_test).", i))


  L <- min(vapply(parts, length, integer(1)))
  parts     <- lapply(parts, function(v) v[seq_len(L)])
  truth_use <- truth[seq_len(L)]
  X_use     <- X_ref[seq_len(L), , drop = FALSE]

  model_names <- names(parts)
  M <- length(model_names)
  
  W <- .gen_weight_grid(M, step = step_w)
  if (nrow(W) == 0) stop("Weight grid is empty, please increase step_w or check M.")

  best <- list(score = -Inf, status = NA_character_, w = rep(NA_real_, M),
               thr = NA_real_, row = NULL, p = NULL)

  if (verbose) cat(sprintf("[Vote-%d] Evaluating %d weight combinations (step=%.2f)...\n", i, nrow(W), step_w))

  for (k in seq_len(nrow(W))) {
    w <- W[k, ]
    P <- 0
    for (j in seq_len(M)) P <- P + as.numeric(parts[[j]]) * w[j]

    pick <- .pick_threshold_constrained_local(truth_use, P, th_grid,
                                              min_spec, min_acc, target_sens)
    r <- pick$best_row
    if (pick$status == "HIT") {
      J <- with(r, Sensitivity + Specificity - 1)
      score <- 1000 + r$Sensitivity + 0.01*J + 1e-4*r$F1   
    } else if (pick$status == "LIMIT") {
      J <- with(r, Sensitivity + Specificity - 1)
      score <- 10 + r$Sensitivity + 0.01*J + 1e-4*r$F1
    } else {
      J <- with(r, Sensitivity + Specificity - 1)
      score <- r$Sensitivity + 0.01*J + 1e-4*r$F1
    }

    if (score > best$score) {
      best <- list(score = score, status = pick$status, w = w,
                   thr = as.numeric(pick$th), row = r, p = P)
      if (verbose && (k %% 50 == 0 || k == nrow(W))) {
        cat(sprintf(" Evaluated %d/%d; Current best: %s  w=%s  thr=%.2f  Se=%.3f Sp=%.3f Acc=%.3f\n",
                    k, nrow(W), pick$status,
                    paste(sprintf("%s=%.2f", model_names, w), collapse=", "),
                    best$thr, r$Sensitivity, r$Specificity, r$Accuracy))
      }
    }
  }

  p_best <- as.numeric(best$p)
  pred   <- factor(ifelse(p_best >= best$thr, "Dyslexia","Normal"),
                   levels = c("Dyslexia","Normal"))
  cnt    <- .cm_counts(truth_use, pred)
  cm_mat <- .cm2matrix(cnt)
  base   <- .metrics_from_counts(cnt)
  aucv   <- suppressWarnings(yardstick::roc_auc_vec(truth_use, p_best, event_level = "first"))

  metrics <- tibble::tibble(
    Dataset = i,
    Status  = best$status,
    Weights = paste(sprintf("%s=%.2f", model_names, best$w), collapse="; "),
    Threshold = round(best$thr, 2),
    AUC = as.numeric(aucv)
  ) |> dplyr::bind_cols(base)

  roc_plot <- {
    roc_obj <- pROC::roc(response = truth_use, predictor = p_best,
                         levels = c("Dyslexia","Normal"), direction = ">")
    pROC::ggroc(roc_obj) +
      ggplot2::geom_abline(linetype = "dashed", colour = "grey50") +
      ggplot2::annotate("text", x = 0.7, y = 0.3,
                        label = paste("AUC =", round(pROC::auc(roc_obj), 3))) +
      ggplot2::labs(title = paste("ROC Curve - VOTE -", i),
                    x = "False Positive Rate", y = "True Positive Rate") +
      ggplot2::theme_bw()
  }

  shap_plot <- NULL; shap_imp_df <- NULL; shap_values <- NULL
  if (top_n_shap > 0) {
    surro <- build_prob_surrogate(X_use, p_best)  
    shap_out <- try({
      compute_shap(
        model         = NULL,
        X_mat         = X_use,
        y_factor      = truth_use,
        pred_fun_prob = function(m, newdata) surro(as.data.frame(newdata)),
        nsim          = 50,
        top_n         = top_n_shap,
        base_use      = "mean_logit"
      )
    }, silent = TRUE)

    if (!inherits(shap_out, "try-error")) {
      shap_plot   <- shap_out$plot        
      shap_imp_df <- shap_out$imp_df
      shap_values <- shap_out$matrix    
    } else if (verbose) {
      message(sprintf("[Vote-%d] SHAP skipped：%s", i, as.character(shap_out)))
    }
  }

  list(
    metrics     = metrics,
    conf_mat    = cm_mat,
    probs_vote  = p_best,
    truth_test  = truth_use,
    threshold   = best$thr,
    weights     = setNames(as.numeric(best$w), model_names),
    oof_bestrow = best$row,
    roc_plot    = roc_plot,
    shap_plot   = shap_plot,
    shap_imp_df = shap_imp_df,
    shap_values = shap_values
  )
}
```

## Run Full-variable Models (Dataset 1)


```{r}

# =====================================================================
# Option A: Train Models from Scratch (Commented out by default)
# You can uncomment the code below to retrain all models.
#
# NOTE ON REPRODUCIBILITY:
# - LR, RF, and XGBoost results are highly stable with the set seed.
# - MLP and Vote models involve Keras backend randomness and multi-threading.
#   Retraining them will result in minor, expected fluctuations in metrics.
# =====================================================================
# # 1. Logistic Regression (LR)
# LR_15sd_results_1 <- run_lr_analysis(
#   train_data = RD_4ML_1_train,
#   test_data  = RD_4ML_1_test,
#   dataset_name = "RD_4ML_1"
# )
# 
# # 2. Random Forest (RF)
# RF_15sd_results_1 <- run_rf_analysis(
#   train_data = RD_4ML_1_train,
#   test_data  = RD_4ML_1_test,
#   dataset_name = "RD_4ML_1"
# )
# 
# # 3. XGBoost (XGB)
# XGB_15sd_results_1 <- run_xgb_analysis(
#   train_data = RD_4ML_1_train,
#   test_data  = RD_4ML_1_test,
#   dataset_name = "RD_4ML_1"
# )

# 4. Multilayer Perceptron (MLP)
# MLP_15sd_results_1 <- run_mlp_analysis(
#   train_data = RD_4ML_1_train,
#   test_data  = RD_4ML_1_test,
#   dataset_name = "RD_4ML_1"
# )

# 5. Voting Ensemble (Vote) 
# res_list_vote <- vote_optimize(
#   i = 1,
#   step_w = 0.10,                       
#   th_grid = seq(0.01, 0.99, 0.01),
#   min_spec = 0.80, 
#   min_acc = 0.80, 
#   target_sens = 0.90,
#   top_n_shap = 10,
#   verbose = TRUE
# )
# =====================================================================
# Option B: Load Pre-trained Models (Recommended)
# - Loading LR, RF, and XGBoost is purely for convenience to save training time.
# - Loading MLP and Vote is REQUIRED to achieve 100% exact reproduction 
#   of the paper's figures, metrics, and DeLong/McNemar test p-values.
# =====================================================================
LR_15sd_results_1  <- readRDS("LR_results.rds")
RF_15sd_results_1  <- readRDS("RF_results.rds")
XGB_15sd_results_1 <- readRDS("XGB_results.rds")
MLP_15sd_results_1 <- readRDS("MLP_results.rds")
res_list_vote      <- readRDS("Vote_results.rds")


```

# full-model results
```{r}
# 1. Extract and summarize metrics for the 5 models
metrics_list <- list(
  LR   = LR_15sd_results_1$metrics,
  RF   = RF_15sd_results_1$metrics,
  XGB  = XGB_15sd_results_1$metrics,
  MLP  = MLP_15sd_results_1$metrics,
  Vote = res_list_vote[[1]]$metrics  
)

# Merge into a dataframe
metrics_tbl_1 <- dplyr::bind_rows(metrics_list, .id = "Algorithm") |>
  dplyr::select(-Threshold_exact, -Status, -Weights)

cat("\n=== Table: Metrics Comparison (Full Variables) ===\n")
print(metrics_tbl_1)

# 2. Sequentially display detailed results for each model
results_list_1 <- list(
  LR   = LR_15sd_results_1,
  RF   = RF_15sd_results_1,
  XGB  = XGB_15sd_results_1,
  MLP  = MLP_15sd_results_1,
  Vote =  res_list_vote[[1]]
)

for (model_name in names(results_list_1)) {
  cat("\n======================================================\n")
  cat("   Model:", model_name, "(Dataset 1: Full Variables)\n")
  cat("======================================================\n")
  
  res <- results_list_1[[model_name]]
  
  # A. Print confusion matrix
  cat("\n--- Confusion Matrix ---\n")
  if (!is.null(res$conf_mat)) {
    cm <- if (inherits(res$conf_mat, "conf_mat")) res$conf_mat$table else res$conf_mat
    print(cm)
  } else {
    cat("No confusion matrix available.\n")
  }
  
  # B. Print specific SHAP values (preview top 10
  if (!is.null(res$shap_values)) {
    cat("\n--- Specific SHAP Values (Top Rows) ---\n")
    print(head(res$shap_values, 10))
  } else if (!is.null(res$shap_data)) {
    cat("\n--- Specific SHAP Data (Top Rows) ---\n")
    print(head(res$shap_data, 10))
  }
  
  # C. Print SHAP
  if (!is.null(res$shap_plot)) {
    cat("\n--- SHAP Plot ---\n")
    print(res$shap_plot)
  }
}
```



## Run XGBoost Sub-models (Dataset 2 to 7)
```{r}
# =====================================================================
# Option A: Train Sub-models from Scratch (Commented out by default)
# XGBoost results are stable. You can uncomment the loop below to retrain 
# the sub-models for Datasets 2 to 7.
# =====================================================================
# for (i in 2:7) {
#   train_name <- paste0("RD_4ML_", i, "_train")
#   test_name  <- paste0("RD_4ML_", i, "_test")
#   data_name  <- paste0("RD_4ML_", i)
#   
#   result_name <- paste0("XGB_15sd_results_", i)
#   
#   assign(result_name, run_xgb_analysis(
#     train_data   = get(train_name),
#     test_data    = get(test_name),
#     dataset_name = data_name
#   ))
# }

# =====================================================================
# Option B: Load Pre-trained XGBoost Sub-models (Recommended)
# Loaded purely for convenience to bypass the time-consuming grid search, 
# allowing immediate reproduction of the sub-model ROC curves and comparisons.
# =====================================================================
XGB_15sd_results_2 <- readRDS("XGB_results_2.rds")
XGB_15sd_results_3 <- readRDS("XGB_results_3.rds")
XGB_15sd_results_4 <- readRDS("XGB_results_4.rds")
XGB_15sd_results_5 <- readRDS("XGB_results_5.rds")
XGB_15sd_results_6 <- readRDS("XGB_results_6.rds")
XGB_15sd_results_7 <- readRDS("XGB_results_7.rds")

```

# XGBoost Sub-models result
```{r}
# 1. Extract and summarize metrics for the 6 sub-models
xgb_sub_metrics <- list()

for (i in 2:7) {
  res_name <- paste0("XGB_15sd_results_", i)
  if (exists(res_name)) {
    res <- get(res_name)
    # Use Sub-model_i as the list name
    xgb_sub_metrics[[paste0("Sub_model_", i)]] <- res$metrics
  }
}

# Combine into a dataframe
metrics_tbl_sub <- dplyr::bind_rows(xgb_sub_metrics, .id = "Domain_Submodel")
cat("\n=== Table: Metrics Comparison (XGBoost Sub-models) ===\n")
print(metrics_tbl_sub)


```



# comparison of full-model
```{r}

truth <- XGB_15sd_results_1$truth_test

p_xgb  <- XGB_15sd_results_1$probs_test
p_vote <- res_list_vote[[1]]$probs_vote

th_xgb  <- XGB_15sd_results_1$threshold
th_vote <- res_list_vote[[1]]$threshold

pred_xgb  <- factor(ifelse(p_xgb >= th_xgb, "Dyslexia", "Normal"), levels = c("Dyslexia", "Normal"))
pred_vote <- factor(ifelse(p_vote >= th_vote, "Dyslexia", "Normal"), levels = c("Dyslexia", "Normal"))


roc_xgb  <- roc(response = truth, predictor = p_xgb, levels = c("Dyslexia", "Normal"), direction = ">", quiet = TRUE)
roc_vote <- roc(response = truth, predictor = p_vote, levels = c("Dyslexia", "Normal"), direction = ">", quiet = TRUE)
delong_res <- roc.test(roc_xgb, roc_vote, method = "delong")
cat(sprintf("AUC DeLong Test p-value: %.4f\n", delong_res$p.value))


acc_table <- table(XGB_Correct = (pred_xgb == truth), VOTE_Correct = (pred_vote == truth))
mcnemar_acc <- mcnemar.test(acc_table)
cat(sprintf("Accuracy McNemar Test p-value: %.4f\n", mcnemar_acc$p.value))


idx_dd <- (truth == "Dyslexia")
sen_table <- table(XGB_Pred = pred_xgb[idx_dd], VOTE_Pred = pred_vote[idx_dd])
mcnemar_sen <- mcnemar.test(sen_table)
cat(sprintf("Sensitivity McNemar Test p-value: %.4f\n", mcnemar_sen$p.value))


idx_td <- (truth == "Normal")
spe_table <- table(XGB_Pred = pred_xgb[idx_td], VOTE_Pred = pred_vote[idx_td])
mcnemar_spe <- mcnemar.test(spe_table)
cat(sprintf("Specificity McNemar Test p-value: %.4f\n", mcnemar_spe$p.value))
```

# comparison of sub-model
```{r}
idx_dd <- (truth == "Dyslexia")
idx_td <- (truth == "Normal")

xgb_models <- list(
  Full_Model = XGB_15sd_results_1,
  Subset_2   = XGB_15sd_results_2,
  Subset_3   = XGB_15sd_results_3,
  Subset_4   = XGB_15sd_results_4,
  Subset_5   = XGB_15sd_results_5,
  Subset_6   = XGB_15sd_results_6,
  Subset_7   = XGB_15sd_results_7
)

p_base    <- xgb_models[[1]]$probs_test
th_base   <- xgb_models[[1]]$threshold
pred_base <- factor(ifelse(p_base >= th_base, "Dyslexia", "Normal"), levels = c("Dyslexia", "Normal"))
roc_base  <- roc(response = truth, predictor = p_base, levels = c("Dyslexia", "Normal"), direction = ">", quiet = TRUE)

comparison_results <- data.frame(
  Subset_Model = character(),
  AUC_Base = numeric(),
  AUC_Sub = numeric(),
  DeLong_p = numeric(),
  Acc_Base = numeric(),
  Acc_Sub = numeric(),
  Acc_McNemar_p = numeric(),
  Sen_Base = numeric(),
  Sen_Sub = numeric(),
  Sen_McNemar_p = numeric(),
  Spe_Base = numeric(),
  Spe_Sub = numeric(),
  Spe_McNemar_p = numeric(),
  stringsAsFactors = FALSE
)

for (i in 2:length(xgb_models)) {
  m_name <- names(xgb_models)[i]
  m_obj  <- xgb_models[[i]]
  
  p_sub    <- m_obj$probs_test
  th_sub   <- m_obj$threshold
  pred_sub <- factor(ifelse(p_sub >= th_sub, "Dyslexia", "Normal"), levels = c("Dyslexia", "Normal"))
  roc_sub  <- roc(response = truth, predictor = p_sub, levels = c("Dyslexia", "Normal"), direction = ">", quiet = TRUE)
  
  delong_test <- roc.test(roc_base, roc_sub, method = "delong")
  
  acc_tab <- table(Base_Correct = (pred_base == truth), Sub_Correct = (pred_sub == truth))
  acc_test <- mcnemar.test(acc_tab)
  
  sen_tab <- table(Base_Pred = pred_base[idx_dd], Sub_Pred = pred_sub[idx_dd])
  sen_test <- mcnemar.test(sen_tab)
  
  spe_tab <- table(Base_Pred = pred_base[idx_td], Sub_Pred = pred_sub[idx_td])
  spe_test <- mcnemar.test(spe_tab)
  
  metrics_sub <- m_obj$metrics
  metrics_base <- xgb_models[[1]]$metrics
  
  comparison_results <- rbind(comparison_results, data.frame(
    Subset_Model  = m_name,
    AUC_Base      = metrics_base$AUC,
    AUC_Sub       = metrics_sub$AUC,
    DeLong_p      = delong_test$p.value,
    
    Acc_Base      = metrics_base$Accuracy,
    Acc_Sub       = metrics_sub$Accuracy,
    Acc_McNemar_p = acc_test$p.value,
    
    Sen_Base      = metrics_base$Sensitivity,
    Sen_Sub       = metrics_sub$Sensitivity,
    Sen_McNemar_p = sen_test$p.value,
    
    Spe_Base      = metrics_base$Specificity,
    Spe_Sub       = metrics_sub$Specificity,
    Spe_McNemar_p = spe_test$p.value
  ))
}

print(comparison_results |> mutate(across(where(is.numeric), ~round(.x, 4))))


truth <- XGB_15sd_results_2$truth_test
idx_dd <- (truth == "Dyslexia")
idx_td <- (truth == "Normal")

sub_models <- list(
  Subset_2 = XGB_15sd_results_2, 
  Subset_3 = XGB_15sd_results_3, 
  Subset_4 = XGB_15sd_results_4, 
  Subset_5 = XGB_15sd_results_5, 
  Subset_6 = XGB_15sd_results_6, 
  Subset_7 = XGB_15sd_results_7 
)

pairs <- combn(names(sub_models), 2, simplify = FALSE)

res_list <- lapply(pairs, function(p) {
  m1_name <- p[1]
  m2_name <- p[2]
  m1 <- sub_models[[m1_name]]
  m2 <- sub_models[[m2_name]]
  
  p1 <- as.numeric(m1$probs_test)
  p2 <- as.numeric(m2$probs_test)
  pred1 <- factor(ifelse(p1 >= m1$threshold, "Dyslexia", "Normal"), levels = c("Dyslexia", "Normal"))
  pred2 <- factor(ifelse(p2 >= m2$threshold, "Dyslexia", "Normal"), levels = c("Dyslexia", "Normal"))
  
  roc1 <- roc(response = truth, predictor = p1, levels = c("Dyslexia", "Normal"), direction = ">", quiet = TRUE)
  roc2 <- roc(response = truth, predictor = p2, levels = c("Dyslexia", "Normal"), direction = ">", quiet = TRUE)
  delong_p <- roc.test(roc1, roc2, method = "delong")$p.value
  
  tab_acc <- table(pred1 == truth, pred2 == truth)
  acc_p <- tryCatch(mcnemar.test(tab_acc)$p.value, error = function(e) NA)
  
  tab_sen <- table(pred1[idx_dd], pred2[idx_dd])
  sen_p <- tryCatch(mcnemar.test(tab_sen)$p.value, error = function(e) NA)
  
  tab_spe <- table(pred1[idx_td], pred2[idx_td])
  spe_p <- tryCatch(mcnemar.test(tab_spe)$p.value, error = function(e) NA)
  
  data.frame(
    Model_A = m1_name,
    Model_B = m2_name,
    AUC_p = delong_p,
    Acc_p = acc_p,
    Sen_p = sen_p,
    Spe_p = spe_p
  )
})

res_df <- bind_rows(res_list)

res_df <- res_df %>%
  mutate(
    AUC_p_adj = p.adjust(AUC_p, method = "holm"),
    Acc_p_adj = p.adjust(Acc_p, method = "holm"),
    Sen_p_adj = p.adjust(Sen_p, method = "holm"),
    Spe_p_adj = p.adjust(Spe_p, method = "holm")
  ) %>%
  select(Model_A, Model_B, 
         AUC_p, AUC_p_adj, 
         Sen_p, Sen_p_adj, 
         Spe_p, Spe_p_adj,
         Acc_p, Acc_p_adj)

print(res_df %>% mutate(across(where(is.numeric), ~round(.x, 4))))
```



# shap picture
```{r}

# =====================================================================
draw_shap_single <- function(model_obj, title_text) {
  if (is.null(model_obj) || is.null(model_obj$shap_plot)) return(NULL)
  
  plot_data <- model_obj$shap_plot$data %>% filter(!is.na(MeanAbsDeltaP))
  
  p <- ggplot(plot_data, aes(x = reorder(Feature, MeanAbsDeltaP), y = MeanAbsDeltaP)) +
    geom_col(width = 0.7, fill = "gray") +  
    geom_text(aes(label = sprintf("%.3f", MeanAbsDeltaP)), 
              hjust = -0.15, size = 3.5, color = "black") +
    coord_flip() +
    labs(title = title_text, x = "Feature", y = "Mean |Δp|") + 
    theme_minimal() +
    scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.15))) +
    theme(
      plot.title = element_text(hjust = 0.5), 
      axis.text.y = element_text(size = 11, face = "bold", color = "black"),
      axis.text.x = element_text(size = 10, color = "black")
    )
  
  return(p)
}

# 1. LR
p_lr <- draw_shap_single(LR_15sd_results_1, "Top 10 SHAP (mean |Δp|) - LR")
if (!is.null(p_lr)) print(p_lr)

# 2. RF
p_rf <- draw_shap_single(RF_15sd_results_1, "Top 10 SHAP (mean |Δp|) - RF")
if (!is.null(p_rf)) print(p_rf)

# 3. XGBoost
p_xgb <- draw_shap_single(XGB_15sd_results_1, "Top 10 SHAP (mean |Δp|) - XGBoost")
if (!is.null(p_xgb)) print(p_xgb)

# 4. MLP
p_mlp <- draw_shap_single(MLP_15sd_results_1, "Top 10 SHAP (mean |Δp|) - MLP")
if (!is.null(p_mlp)) print(p_mlp)

# 5. Vote 
p_vote <- draw_shap_single(res_list_vote[[1]], "Top 10 SHAP (mean |Δp|) - Vote")
if (!is.null(p_vote)) print(p_vote)
```



# ROC
```{r}
get_roc_data <- function(model_obj, model_name) {
  
  truth <- model_obj$truth_test
  
  if (!is.null(model_obj$probs_test)) {
    prob <- model_obj$probs_test
  } else if (!is.null(model_obj$probs_vote)) {
    prob <- model_obj$probs_vote
  } else {
    stop(sprintf("Error: No probability data found in model [%s]!", model_name))
  }
  
  if(is.data.frame(truth)) truth <- truth[[1]]
  if(is.list(truth)) truth <- unlist(truth)
  
  if(is.data.frame(prob)) prob <- prob[[ncol(prob)]]
  if(is.list(prob)) prob <- unlist(prob)
  
  prob <- as.numeric(prob)
  
  my_roc <- roc(response = truth, predictor = prob, quiet = TRUE)
  auc_val <- as.numeric(auc(my_roc))
  
  data.frame(
    FPR = 1 - my_roc$specificities,
    TPR = my_roc$sensitivities,
    Model = sprintf("%s (AUC = %.3f)", model_name, auc_val)
  )
}

my_colors <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#A65628", "#F781BF")


roc_df_1 <- bind_rows(
  get_roc_data(LR_15sd_results_1,  "LR"),
  get_roc_data(RF_15sd_results_1,  "RF"),
  get_roc_data(XGB_15sd_results_1, "XG-Boost"),
  get_roc_data(MLP_15sd_results_1, "MLP"),
  get_roc_data(res_list_vote[[1]], "Vote Ensemble") 
) %>%
  mutate(Model = factor(Model, levels = unique(Model)))

p_roc_1 <- ggplot(roc_df_1, aes(x = FPR, y = TPR, color = Model)) +
  geom_path(linewidth = 1.2) + 
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray50", linewidth = 0.8) + 
  scale_color_manual(values = my_colors[1:5]) + 
  labs(title = "ROC Curves", x = "False Positive Rate (1 - Specificity)", y = "True Positive Rate (Sensitivity)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title = element_text(face = "plain", size = 12),
    axis.text = element_text(size = 11, color = "black"),
    legend.position = c(0.75, 0.25), 
    legend.background = element_rect(fill = "white", color = "black", linewidth = 0.5),
    legend.title = element_blank(),
    legend.text = element_text(size = 10)
  )
print(p_roc_1)


roc_df_2 <- bind_rows(
  get_roc_data(XGB_15sd_results_2, "Whole-Character Phonology"),
  get_roc_data(XGB_15sd_results_3, "Whole-Character Orthography"),
  get_roc_data(XGB_15sd_results_4, "Whole-Character Semantics"),
  get_roc_data(XGB_15sd_results_5, "Sub-lexical Phonology"),
  get_roc_data(XGB_15sd_results_6, "Sub-lexical Orthography"),
  get_roc_data(XGB_15sd_results_7, "Sub-lexical Semantics")
) %>%
  mutate(Model = factor(Model, levels = unique(Model)))

p_roc_2 <- ggplot(roc_df_2, aes(x = FPR, y = TPR, color = Model)) +
  geom_path(linewidth = 1.2) + 
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray50", linewidth = 0.8) + 
  scale_color_manual(values = my_colors[1:6]) + 
  labs(title = "ROC Curves", x = "False Positive Rate (1 - Specificity)", y = "True Positive Rate (Sensitivity)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title = element_text(face = "plain", size = 12),
    axis.text = element_text(size = 11, color = "black"),
    legend.position = c(0.75, 0.25), 
    legend.background = element_rect(fill = "white", color = "black", linewidth = 0.5),
    legend.title = element_blank(),
    legend.text = element_text(size = 10)
  )

print(p_roc_2)
```


